The Practical Application of Computerised Multistage Tests

### 2021-03-05 article.Rmd
Use better test data

---
title: "The Practical Application of Computerised Multistage Tests"
description: A practical guide to the design and development of computerised multistage
  tests.
author:
- name: James Page
  affiliation: SP Educational Consulting
  affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
version: 1
output:
  pdf_document:
    highlight: pygments
    latex_engine: lualatex
    toc: true
    toc_depth: 2
    fig_caption: true
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    fig_caption: true
  word_document:
    toc: true
    toc_depth: 2
    fig_caption: true
bibliography: references.bib
---

@@

@@

## IRT for Mulitstage Test Design

IRT is used to inform calibration of the test items after they've been developed by item writers, as well as construction of the modules and routing between them. Scoring of the test can be derived from the ability estimates. Typically the initial module (known as the routing module) will have an overall difficulty of 0 (ie. a mix of easy and hard items), with subsequent routing directing candidates to easier or more difficult modules in order to 'hone in' on an accurate $\hat{theta}$ (ability estimate) at the end of the test.

A typical modular design for a MST will look like this (Figure 3):

@@

This is an example of a 1-2-4 design, but others are possible. Important questions to answer during test design are what module topology to use (the number of stages and modules per stage), and what should be the overall difficulty levels for each module (represented as the mean item difficulty, $\bar{b}$).

@@

In this part we describe the steps needed in the development of MSTs. Here we turn our attention to initial validation to confirm the soundness of the test design that was conceived at the first stage of test development described in Part 1 of this article.

@@

In the pre-trial stage, it is advisable to run simulations to help us answer questions raised in Part 1 of this article, and inform final decisions about test design. It's always less expensive to make modifications to a design early on in the process, before significant investment in item writing and trialling has been made. The ability to simulate a test at the design phase can be used to understand the constraints of the design at an early stage and make changes in a low-cost, low-risk way. In this section we'll show how the `mstR` R package can be used to run simulations for a MST design. First we'll discuss the data sources that we'll be using.

@@

In order to perform a simulated test, there are a number of different data sources that are required.

The first class of these data sources is the one that is used to store item information. The key entities to store here:

@@

Finally, test response data is required. Essentially this is a matrix containing rows of candidates and columns of items. Cells will contain a 1 (item answered correctly) or a 0 (item answered incorrectly). Items that were not presented (for example, from a module that was not in the test path for a candidate) are left blank and will be classed as NA.

This data may be stored as flat files (in CSV format for instance), or held in database tables, or in a mix of different sources. Typically you would want to keep them in a cloud storage system (for example in AWS S3 buckets or an RDS database) to ensure that the data is stored in a way that is secure, accessible and geographically distributed for resilience.

@@

An initial simulation of your test to experiment with different designs can be done by using the `mstR` library (see @magis2017computerized). Here the aim is to select a set of designs and see if there is an optimal design. It will also help understand how the scoring of the test should be implemented, including the number correct (NC) scores that will be used for routing, and the ability estimates for final scoring and reporting.

@@

The aim here is to sift several different test designs; and re-run simulations easily.

@@

In addition to the routing paths, a set of cut-off criteria must be given. These are expressed in terms of $\hat{theta}$ using the `cutoff` matrix. To use an example here, the rule `cutoff[module('B'),] <- c(-Inf, 0.0)` means that candidates with ability estimates from minus infinity to zero will be routed to module B. The rule `cutoff[module('C'),] <- c(0.0, Inf)` means that candidates with ability estimates from zero to plus infinity will be routed to module C.

@@

```{r show simulated test, echo=FALSE}
plot(dfMSTResponse)
```

@@

```{r test-design-params, echo=FALSE}
designParams <- read_csv('mst-designs.csv', skip = 1) %>%
  clean_names() %>%
  dplyr::select(-design, -total)

@@

```{r MST designs}
mstDesign <- function(stage1, stage2, stage3,
                      module_a, module_b, module_c, module_d, module_e, module_f, module_g) {
  matrixN <- stage1 + (2 * stage2) + (4 * stage3)
  modules <- matrix(0, matrixN, 7)
  moduleA_items <- stage1
  moduleB_items <- moduleA_items + stage2
  moduleC_items <- moduleB_items + stage2
  moduleD_items <- moduleC_items + stage3
  moduleE_items <- moduleD_items + stage3
  moduleF_items <- moduleE_items + stage3
  moduleG_items <- moduleF_items + stage3

@@

  return(list("modules" = modules, "itemBank" = itembank))
}

@@

The results can now be graphed. Following @ippel2020efficient and @magis2017computerized we compute the averaged signed bias (ASB) and root mean squared error (RMSE) values for each design and for each ability level. ASB is the mean difference between estimated and true ability levels. RMSE is the square root of the mean of the squared differences between estimated and true ability levels. With both measures, lower is better.

@@

```{r simultation graphs}
ggplot(data = dfASB, aes(x = theta)) +
  geom_line(aes(y = X1)) + 
  geom_line(aes(y = X2)) +
  geom_line(aes(y = X3)) +
  geom_line(aes(y = X4)) +
  geom_line(aes(y = X5)) +
  geom_line(aes(y = X6)) +
  geom_hline(yintercept = 0.0, linetype="dotted") +
  labs(x = "theta", y = "ASB")

ggplot(data = dfRMSE, aes(x = theta)) +
  geom_line(aes(y = X1)) + 
  geom_line(aes(y = X2)) +
  geom_line(aes(y = X3)) +
  geom_line(aes(y = X4)) +
  geom_line(aes(y = X5)) +
  geom_line(aes(y = X6)) +
  labs(x = "theta", y = "RMSE")
```

@@

| Design | Stage 1 | Stage 2 | Stage 3 | Total |
| ------ | ------: | ------: | ------: | ----: |
| 1      | 12      | 8       | 10      | 30    |

@@

head(dfTestResponses[1:8])
```

@@

(dfCronbachsAlpha)

@@

You can see in Figure 11 that some of the outlier items have a very large divergence from the mean; it is best to remove these items from the item bank as they have too large an error to be used in a live test.

@@

Once difficulty values have been calculated by fitting the Rasch model, the item bank should be updated to reflect this. The item difficulty ($b$ parameter) is a key component of the item, and should be stored along with the other information about the item (including its rated CEFR level by the subject expert panelists). If a 2PL model is being used, then the discrimination ($a$ parameter) would also be stored.

@@

Something that we see here that is odd is that our module E (easy) is actually a little bit more difficult than our module F (hard). Also all of the modules are quite similar in terms of difficulty (the MCCs are quite close to each other), and moreover the lines are quite flat - there is little discrimination between them. Again we may want to look at moving some items between modules, or rejecting items with poor discrimination,

>>>>

---
title: "The Practical Application of Computerised Multistage Tests"
description: A practical guide to the design and development of computerised multistage
  tests.
author:
- name: James Page
  affiliation: SP Educational Consulting
  affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
version: 1
output:
  pdf_document: 
    highlight: tango
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
    fig_caption: yes
    fig_width: 5
    fig_height: 4
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    fig_caption: true
  word_document:
    toc: true
    toc_depth: 2
    fig_caption: true
bibliography: references.bib
---

@@

See Chapter 1 of @lane2015handbook for a concise introduction to the main considerations of developing a new test.

@@

## IRT for Multistage Test Design

IRT is used to inform calibration of the test items after they've been developed by item writers, as well as construction of the modules and routing between them. Scoring of the test can be derived from the ability estimates. Typically the initial module (known as the routing module) will have an overall difficulty of 0 (ie. a mix of easy and hard items), with subsequent routing directing candidates to easier or more difficult modules in order to 'hone in' on an accurate $\hat{\theta}$ (ability estimate) at the end of the test.

A typical modular design for a MST will look like the one in Figure 3 which is an example of a 1-2-4 design. Other designs are possible.

@@

Important questions to answer during test design include what module topology to use (the number of stages, and the number of modules per stage), and what overall difficulty level should be used for each module (represented by the mean item difficulty, $\bar{b}$).

@@

In this part we describe the steps needed for the development of MSTs. Here we turn our attention to initial validation of a test design.

@@

In the pre-trial stage, it is useful to run simulations to help us answer the questions posed in the previous part of this article. The results of these simulations will help inform final decisions about the test design. It's always less expensive to make modifications to a design early on in the process, before significant investment in item writing and trialling has been made. The ability to simulate a test at the design phase can be used to understand the constraints of the design at an early stage and make changes in a low-cost, low-risk way. In this section we'll show how the `mstR` R package can be used to run simulations for a MST design. First though we'll cover the data sources that we'll be using.

@@

In order to perform a simulated test, there are a number of different data sources that are required. To start with we need to consider the following entities that will be used to construct a test:

@@

Finally, test response data is required. Essentially this is a matrix containing rows of candidates and columns of items. Cells in the matrix contain a 1 (item answered correctly) or a 0 (item answered incorrectly). Items that were not presented (for example, from a module that was not in the test path for a candidate) are left blank and are classed as NA.

This data may be stored as flat files (in CSV format for instance), or held in database tables, or in a mix of different sources. Typically a cloud storage system will be used (for example in AWS S3 buckets or in a cloud database such as Amazon RDS). This is to ensure that the data is stored in a way that is secure, accessible and geographically distributed (for resilience).

@@

An initial simulation of the test can be used to experiment with different designs. We show how to do this using the `mstR` library (see @magis2017computerized). Here the aim is to select a set of designs and see if there is an optimal design. It will also help understand how the scoring of the test should be implemented, including the number correct (NC) scores that will be used for routing, and the ability estimates for final scoring and reporting.

@@

The aim here is to sift several different test designs with a view of converging on the optimal design for the . The ability to re-run simulations easily with different design parameters provides a powerful means of trying out different design ideas before incurring the costs of engaging item writers, raters and trial candidates.

@@

In addition to the routing paths, a set of cut-off criteria must be given. These are expressed in terms of $\hat{\theta}$ using the `cutoff` matrix (see below). To use an example here, the rule `cutoff[module('B'),] <- c(-Inf, 0.0)` means that candidates with ability estimates from minus infinity to zero will be routed to module B. The rule `cutoff[module('C'),] <- c(0.0, Inf)` means that candidates with ability estimates from zero to positive infinity will be routed to module C.

@@

```{r show simulated test, echo=FALSE, fig.cap="Example module routing"}
plot(dfMSTResponse)
```

@@

```{r test-design-params, echo=FALSE}
designParams <- read_csv('mst-designs.csv', skip = 1) %>%
  clean_names() %>%
  dplyr::select(-total)

@@

```{r MST designs}
mstDesign <- function(design, stage1, stage2, stage3,
                      module_a, module_b, module_c, module_d, module_e, module_f, module_g) {
  matrixN <- stage1 + (2 * stage2) + (4 * stage3)
  modules <- matrix(0, matrixN, 7)
  moduleA_items <- stage1
  moduleB_items <- moduleA_items + stage2
  moduleC_items <- moduleB_items + stage2
  moduleD_items <- moduleC_items + stage3
  moduleE_items <- moduleD_items + stage3
  moduleF_items <- moduleE_items + stage3
  moduleG_items <- moduleF_items + stage3

@@

  return(list("design" = design, "modules" = modules, "itemBank" = itembank))
}

@@

The results can now be graphed. Following @ippel2020efficient, and @magis2017computerized we compute the averaged signed bias (ASB) and root mean squared error (RMSE) values for each design and for each ability level. ASB is the mean difference between estimated and true ability levels. RMSE is the square root of the mean of the squared differences between estimated and true ability levels. With both measures, lower is better.

@@

```{r ASB, fig.cap="ASB"}
ggplot(data = dfASB, aes(x = theta)) +
  geom_line(aes(y = X1)) + 
  geom_line(aes(y = X2)) +
  geom_line(aes(y = X3)) +
  geom_line(aes(y = X4)) +
  geom_line(aes(y = X5)) +
  geom_line(aes(y = X6)) +
  geom_hline(yintercept = 0.0, linetype="dotted") +
  labs(x = "theta", y = "ASB")
```

```{r RMSE, fig.cap="RMSE"}
ggplot(data = dfRMSE, aes(x = theta)) +
  geom_line(aes(y = X1)) + 
  geom_line(aes(y = X2)) +
  geom_line(aes(y = X3)) +
  geom_line(aes(y = X4)) +
  geom_line(aes(y = X5)) +
  geom_line(aes(y = X6)) +
  labs(x = "theta", y = "RMSE")
```

@@

```{r}
knitr::kable(
  head(designParams, n = 1), booktabs = TRUE,
  caption = 'MST Design 1'
)
```

@@

knitr::kable(
  head(dfTestResponses[1:8]), booktabs = TRUE,
  caption = 'Example Test Reponses'
)
```

@@

knitr::kable(
  head(dfCronbachsAlpha), booktabs = TRUE,
  caption = "Cronbach's Alpha"
)

@@

This chart makes it easy to spot any outlier items have a very large divergence from the mean; it is best to remove these items from the item bank as they have too large an error to be used in a live test. Usually this is the result of them having all 1's or 0's (in other words they are too easy or too difficult for all candidates who took them).

@@

Once difficulty values have been calculated by fitting the Rasch model, the item bank should be updated to reflect this. The item difficulty ($b$ parameter) is a key component of the item, and should be stored along with the other information about the item (including its rated CEFR level by the subject expert panellists). If a 2PL model is being used, then the discrimination ($a$ parameter) would also be stored.

@@

All of the modules are quite similar in terms of difficulty (the MCCs are quite close to each other), and moreover the lines are quite flat - there is little discrimination between them. So we may want to look at moving some items between modules, or rejecting items with poor discrimination,
